---
title: "Assignment_2_LP_solve"
author: "Hitaishi Bairapureddy"
date: "2024-09-20"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question : 

The Weigelt Corporation has three branch plants with excess production capacity. Fortunately,
the corporation has a new product ready to begin production, and all three plants have this
capability, so some of the excess capacity can be used in this way. This product can be made in
three sizes--large, medium, and small--that yield a net unit profit of $420, $360, and $300,
respectively. Plants 1, 2, and 3 have the excess capacity to produce 750, 900, and 450 units per
day of this product, respectively, regardless of the size or combination of sizes involved.
The amount of available in-process storage space also imposes a limitation on the production
rates of the new product. Plants 1, 2, and 3 have 13,000, 12,000, and 5,000 square feet,
respectively, of in-process storage space available for a day's production of this product. Each
unit of the large, medium, and small sizes produced per day requires 20, 15, and 12 square feet,
respectively.
Sales forecasts indicate that if available, 900, 1,200, and 750 units of the large, medium, and
small sizes, respectively, would be sold per day.
At each plant, some employees will need to be laid off unless most of the plant’s excess
production capacity can be used to produce the new product. To avoid layoffs if possible,
management has decided that the plants should use the same percentage of their excess capacity
to produce the new product.
Management wishes to know how much of each of the sizes should be produced by each of the
plants to maximize profit


```{r}
WC<- matrix(c(900,"$420",20,1200, "$360",15,750,"$300",12),ncol =3,byrow=TRUE)
# Specifying the column names and row names
colnames(WC)<-c("Sales","Profit","Material")
rownames(WC)<-c("Large","Medium","Small")
table1=as.table(WC)
table1
```

 + Objective function:
 
            MaxZ=420(A1+A2+A3)+360(B1+B2+B3)+300(C1+C2+C3)

 Also given as,  MaxZ=420A1+360B1+300C1+420A2+360B2+300C2+420A3+360B3+300C3

                     A1+B1+C1<=750

                     A2+B2+C2<=900

                     A3+B3+C3<=450

                20A1+15B1+12C1<=13000

                20A2+15B2+12C2<=12000

                20A3+15B3+12C3<=5000

                   A1+A2+A3<=900

                   B1+B2+B3<=1200

                  C1+C2+C3<=750

Non Negativity Constraints

                    A1,A2,A3,B1,B2,B3,C1,C2,C3>=0

                 A1+B1+C1/750 = A2+B2+C2/900 = A3+B3+C3/450

             900A1+900B1+900C1−750A2−750B2−750C2+0A3+0B3+0C3=0

              0A1+0B1+0C1+450A2+450B2+450C2−900A3−900B3−900C3=0

              450A1+450B1+450C1+0A2+0B2+0C2−750A3−750B3−750C3=0




```{r}
#install.packages("lpSolve")
library("lpSolve")
```


```{r}
#Defining objective function
objective_fn <- c(420,360,300,420,360,300,420,360,300)
```

```{r}
#Defining the constraint function
constraint_fn <- matrix(c(1,1,1,0,0,0,0,0,0,
                       0,0,0,1,1,1,0,0,0,
                       0,0,0,0,0,0,1,1,1,
                       20,15,12,0,0,0,0,0,0,
                       0,0,0,20,15,12,0,0,0,
                       0,0,0,0,0,0,20,15,12,
                       1,0,0,1,0,0,1,0,0,
                       0,1,0,0,1,0,0,1,0,
                       0,0,1,0,0,1,0,0,1,
                       900,900,900,-750,-750,-750,0,0,0,
                       0,0,0,450,450,450,-900,-900,-900,
                       450,450,450,0,0,0,-750,-750,-750),nrow = 12,byrow = TRUE)
constraint_fn
```

```{r}
# Direction of Constraints
con_direction <- c('<=',

           '<=',
           
           '<=',
           
           '<=',
           
           '<=',
           
           '<=',
           
           '<=',
           
           '<=',
           
           '<=',
           '=',
           '=',
           '=')

#Determining the constants that means the right hand side values - rhs.fn


rhv.fn <- c(750,
         900,
         450,
         13000,
         12000,
         5000,
         900,
         1200,
         750,
         0,
         0,
         0)

          
#Solving the problem using lp function basing the objective function that means to maximize the profit

maxi_cost=lp('max',objective_fn, constraint_fn, con_direction, rhv.fn)

maxi_cost
```

```{r}
maxi_cost$solution
maxi_cost$solution[1]
maxi_cost$solution[2]
maxi_cost$solution[3]
maxi_cost$solution[4]
maxi_cost$solution[5]
maxi_cost$solution[6]
maxi_cost$solution[7]
maxi_cost$solution[8]
maxi_cost$solution[9]
```


