---
title: "Assignment3_QMM"
author: "Hitaishi Bairapureddy"
date: "2024-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Question:

 +  Heart Start produces automated external defibrillators (AEDs) in each of two different plants (A
and B). The unit production costs and monthly production capacity of the two plants are
indicated in the table below. The AEDs are sold through three wholesalers. The shipping cost
from each plant to the warehouse of each wholesaler along with the monthly demand from each
wholesaler are also indicated in the table. How many AEDs should be produced in each plant,
and how should they be distributed to each of the three wholesaler warehouses so as to minimize
the combined cost of production and shipping?


```{r}
# Load lpSolve package
#install.packages("lpSolve")
library(lpSolve)
```


```{r}
# Creating a dataframe for the transportation problem
transport_data <- data.frame(Warehouse_A = c("$22", "$16", 80),
                              Warehouse_B = c("$14", "$20", 60),
                              Warehouse_C = c("$30", "$24", 70),
                              Production_Cost_Per_Unit = c("$600", "$625", " "),
                              Monthly_Capacity = c(100, 120, " "))
rownames(transport_data) <- c("Plant 1", "Plant 2", "Total Demand")

# Generating a formatted table
library(knitr)
# install.packages("kableExtra")
library(kableExtra)

transport_data %>%
  kable() %>%
  kable_classic(full_width = FALSE, html_font = "Arial") %>%
  column_spec(1, border_right = TRUE, bold = TRUE, color = "white", background = "#2c3e50") %>%
  column_spec(2:4, background = "#ecf0f1") %>%
  column_spec(4, border_right = TRUE) %>%
  column_spec(5:6, background = "#e8f8f5") %>%
  row_spec(2, extra_css = "border-bottom: 2px solid #34495e") %>%
  row_spec(0, bold = TRUE, color = "white", background = "#3498db") %>%
  add_header_above(header = c(" " = 1, "Unit Shipping Cost" = 3, " " = 2), 
                   color = "white", background = "#2980b9")

```

## STEPS

### 1. Setup of Decision Variables

Let:
- **\(x_{A1}\)**: units shipped from Plant A to Warehouse 1
- **\(x_{A2}\)**: units shipped from Plant A to Warehouse 2
- **\(x_{A3}\)**: units shipped from Plant A to Warehouse 3
- **\(x_{B1}\)**: units shipped from Plant B to Warehouse 1
- **\(x_{B2}\)**: units shipped from Plant B to Warehouse 2
- **\(x_{B3}\)**: units shipped from Plant B to Warehouse 3

### 2. Objective Function

The total cost \(Z\) consists of both production and shipping costs. The combined cost function to minimize is given by:

\[
Z = 600(x_{A1} + x_{A2} + x_{A3}) + 625(x_{B1} + x_{B2} + x_{B3}) + 22x_{A1} + 14x_{A2} + 30x_{A3} + 16x_{B1} + 20x_{B2} + 24x_{B3}
\]

### 3. Constraints

#### Demand Constraints:
- For Warehouse 1:
\[
x_{A1} + x_{B1} = 80
\]
- For Warehouse 2:
\[
x_{A2} + x_{B2} = 60
\]
- For Warehouse 3:
\[
x_{A3} + x_{B3} = 70
\]

#### Supply Constraints:
- Plant A's capacity:
\[
x_{A1} + x_{A2} + x_{A3} \leq 100
\]
- Plant B's capacity:
\[
x_{B1} + x_{B2} + x_{B3} \leq 120
\]

#### Non-negativity Constraints:
\[
x_{A1}, x_{A2}, x_{A3}, x_{B1}, x_{B2}, x_{B3} \geq 0
\]

## Mathematical Formulation

We aim to minimize the total production and shipping cost, subject to the demand and supply constraints.



Minimize:

\[
Z = 600(x_{A1} + x_{A2} + x_{A3}) + 625(x_{B1} + x_{B2} + x_{B3}) + 22x_{A1} + 14x_{A2} + 30x_{A3} + 16x_{B1} + 20x_{B2} + 24x_{B3}
\]

### Subject to:

- Demand Constraints:

\[
x_{A1} + x_{B1} = 80
\]
\[
x_{A2} + x_{B2} = 60
\]
\[
x_{A3} + x_{B3} = 70
\]

- Supply Constraints:

\[
x_{A1} + x_{A2} + x_{A3} \leq 100
\]
\[
x_{B1} + x_{B2} + x_{B3} \leq 120
\]

- Non-Negativity Constraints:

\[
x_{A1}, x_{A2}, x_{A3}, x_{B1}, x_{B2}, x_{B3} \geq 0
\]

This formulation includes all the necessary constraints and objective function for the transportation problem. Next, we can use this model to solve for the optimal values using linear programming methods.


```{r}
# Specifying the costs related to production and shipping.
# Order of decision variables:
# [Plant 1 to Warehouse A, plant 1 to Warehouse B, plant 1 to Warehouse C, 
#  Plant 2 to Warehouse A, plant 2 to Warehouse B, plant 2 to Warehouse C]
production_shipping_costs <- c(600 + 22, 600 + 14, 600 + 30, 
                               625 + 16, 625 + 20, 625 + 24)

```


```{r}
# Establishing the matrix to represent all constraints for the transportation problem.
# Each row corresponds to a specific constraint:
# 1) Capacity of plant 1, 2) Capacity of plant 2
# 3) Demand at Warehouse A, 4) Demand at Warehouse B, 5) Demand at Warehouse C
constraint_matrix <- matrix(c(1, 1, 1, 0, 0, 0,   # Capacity for Plant 1
                              0, 0, 0, 1, 1, 1,   # Capacity for Plant 2
                              1, 0, 0, 1, 0, 0,   # Demand for Warehouse A
                              0, 1, 0, 0, 1, 0,   # Demand for Warehouse B
                              0, 0, 1, 0, 0, 1),  # Demand for Warehouse C
                            nrow = 5, byrow = TRUE)

```



```{r}
# Specifying the right-hand side values for the constraints.
rhs_values <- c(100, 120, 80, 60, 70)  # Capacities and Demands

```



```{r}
# Defining the relationship types for the constraints
# "less than or equal to" for factory capacities, "equal to" for warehouse demands
constraint_directions <- c("<=", "<=", "=", "=", "=")

```

```{r}
# Solving the transportation problem using the lp() function
transport_solution <- lp("min", production_shipping_costs, constraint_matrix, constraint_directions, rhs_values)
transport_solution

```

```{r}
# Outputting the results
if (transport_solution$status == 0) 
  # If an optimal solution is found then print the results
  print("Optimal solution found:")
  transport_solution$solution_matrix <- matrix(transport_solution$solution, nrow = 2, byrow = TRUE)
  rownames(transport_solution$solution_matrix) <- c("Plant A", "Plant B")
  colnames(transport_solution$solution_matrix) <- c("Warehouse 1", "Warehouse 2", "Warehouse 3")
  print(transport_solution$solution_matrix)


```

```{r}
# Print the optimal solution (how much to ship from each plant(plant 1 and plant 2) to each warehouse)
print("Optimal number of AEDs shipped from each plant to each warehouse:")
transport_solution$solution
```



```{r}
# Printing the minimized total cost without the dollar sign
print(paste("Minimized total cost value: ", transport_solution$objval))

```

## Conclusion:

The optimal solution demonstrates an efficient allocation of production and distribution that minimizes costs while meeting all demand and capacity constraints. By utilizing Plant A at full capacity and Plant B slightly under capacity, Heart Start can effectively meet the demand of all three warehouses.


### Key findings:

Plant A should produce 100 units (at full capacity):

60 units shipped to Warehouse B
40 units shipped to Warehouse C


Plant B should produce 110 units (10 under capacity):

80 units shipped to Warehouse A
30 units shipped to Warehouse C


The minimized total cost is $132,790.


This solution highlights the importance of strategic planning in supply chain management. It shows that the lowest-cost option isn't always to produce at the plant with the lowest production cost, as shipping costs play a crucial role in the overall expense.
